---
title: "Automated report example"
author: "Author"
date: "March 2021"
output:
  pdf_document: default
  word_document: default
  html_document: default
params:
  year:
    label: Year
    value: 2017
    input: slider
    min: 2010
    max: 2021
    step: 1
    sep: ''
  data:
    label: 'Input dataset:'
    value: some_data.csv
    input: file
---

\vspace{0.3cm}

```{r}

# the parameters in the Rmd-file header can be chosen interactively, 
# by using "Knit with parameters"

# or directly, by running:

# rmarkdown::render("RvalPractice0.Rmd", params = list(
#   year = 2021,
#   file = "our_favourite.csv"
# ))

# Try this!

```

# Useful links:

- for downloading and installing R, RStudio on MacOSX, see the example at:
https://web.stanford.edu/~kjytay/courses/stats32-aut2018/Session%201/Installation%20for%20Mac.html 

- for using open data from the web:
https://theodi.org/article/how-to-use-r-to-access-data-on-the-web/

(_hint_: use functions like read.csv(), read.url(), various API's or the RCurl-package )



# Important notes:

- We may all raise issues in the github repository https://github.com/violetacln/learnRval  
about any questions, proposals, ideas we might like to share!


- We do as much as possible here, in our interactive meeting: 

  - we run the examples included in these RvalPractice - files

  - we could also propose new ones and/or run code on different data.



# Introduction: 

## Data-set and main information needed for validation

about: time-variables, modeled - variables, imputed-variables, if the case

Assume a data-frame has been built.

Assume rules set has been built.

Example of notations: 
  
  ```{r datadef, echo=TRUE, eval=FALSE}
  
## data rules, built, read into a data frame and then a validator object
# vrules <- 

## modifier rules, built, read into a data frame and then a validator object
# mrules <- 

## main data set as a data.frame
#df <- 

##data sets which need to be compared (could be same data measured at 
##different moments in time), as data.frames
#df1<- 
#df2<-

##time series we need to check, if it is ths case
#tsuniv
#tsmultiv
```

__Important note:  when we want to create our report and show all results, we make "eval=TRUE" in the R-chunks which follow.__





# Part 1: Data overview and examples of validation rules


## Simplest data overview and univariate analysis

Let us try this example for our favorite data set (df).

```{r, eval=FALSE, echo=TRUE}
library(validate)

df <- SBS2000   # or iris, or any real data set you like!
##variable types
dnames <-names( DataExplorer::split_columns(df)$discrete)
cnames <- names( DataExplorer::split_columns(df)$continuous)

##short overview
  #summary(df)
  Hmisc::describe(df)
  
##data univariate plots
  DataExplorer::plot_intro(df)
  DataExplorer::plot_missing(df)
  funModeling::freq(df)  #DataExplorer::plot_bar(df, maxcat=450)
  funModeling::plot_num(df) #DataExplorer::plot_histogram(df)
  
##qq-plots, overall and conditional
  DataExplorer::plot_qq(df)
  DataExplorer::plot_qq(df, by="size")   ## Species, if use df <- iris

```


 - Let us repeat for data set "diamonds" in "ggplot2" package!

 - We can also try other functions/packages which show same type of overview.
Which ones did/do you prefer?



## Error Detection: Confronting data and rules

See exercise 1 in Part 1 - course slides:

(propose/run F, Cond1, Cond2, forbidden_keys, is_complete, for SBS2000 or samplonomy data)


### ex.1.1: Examples of "balance" conditions

 - First, from https://cran.r-project.org/web/packages/validate/vignettes/ 


```{r , echo=FALSE, eval=FALSE}  

# no echo here! this just for your convenience, do a run-test

library(validate)
data("SBS2000")
str(SBS2000)

rules1 <- validator(
    total.rev - profit == total.costs
  , turnover + other.rev == total.rev
  , profit <= 0.6*total.rev
)

out1 <- confront(SBS2000, rules1)
summary(out1)
plot(out1)

```




 - Second example, built by us, like:


```{r , echo=TRUE, eval=FALSE}
#rule_ex1 <- validator(F>= 0)

data(diamonds, package="ggplot2")
str(diamonds)

rules1B <- validator(
   price >300
  , depth * carat > 11
  , depth * carat < 328  # change this value into lower/higher and see what happens!
)

out1B <- confront(diamonds, rule1s1B)
summary(out1B)

```


### ex.1.2. Example of conditional restrictions (If Condition 1, then Condition 2) 


```{r , echo=TRUE, eval=FALSE}
# rule_ex2 <- validator(if (Cond1) Cond2)

# plot(diamonds$carat, diamonds$price)

rules2 <- validator(if (carat > 4) price > 1500)
out2 <- confront(diamonds, rules2)
summary(out2)

```



### ex.1.3. Forbidden patterns

```{r, echo=TRUE, eval=FALSE}
# rule_ex3 <- validator( does_not_contain(glob(forbidden_keys))) )

# data frame of combinations we do not allow
our_forbidden <- data.frame(cut="Fa*", clarity="VVS*")
  
rules3 <- validator(does_not_contain(glob(forbidden_keys=our_forbidden)))
out3 <- confront(diamonds, rules3)

summary(out3)

# violating(diamonds, out3)

```


### ex.1.4. Completeness of records

```{r, echo=TRUE, eval=FALSE}
# is_complete($Y_1$, ...)  # result: one T/F per line
# all_complete(Y)         # result: one T/F

rules4 <- validator(is_complete(depth))
out4 <- confront(diamonds, rules4)
summary(out4)

# introduce some NA values, using naniar -  package
something <- max(diamonds$x)
diamonds4b <- naniar::replace_with_na(diamonds, replace=list(x = something) )
#or 
# we might like to use dplyr - package:
# diamonds4b <- diamonds
# diamonds4b$x <- dplyr::na_if(diamonds4b$x, something)

rules4b <- validator( is_complete(x)
                    , all_complete(x)
                    , is_complete(x,y,z)
                    )
out4b <- confront(diamonds4b, rules4b)
summary(out4b)

# note the differences!

```



### ex. 1.5. More checks = more functions!

Examples of:

- Conditions on statistical summary measures (per variable(s))

```{r, echo=TRUE, eval=FALSE}

# vs1 example in the slides
rules5 <- validator(
                 mean(x) > 5
                , cor(price, carat) > 0
                , range(price)[1] < 500
                , sd(y) > 1
                    )
out5 <- confront(diamonds, rules5)
summary(out5)

```



- Rules for aggregated data: _rule(Cond,by)_ 


```{r, echo=TRUE, eval=FALSE}

#vs2 example in our slides
rules6 <- validator( 
              exists_one(color=="J", by=list(cut, carat))
            , exists_any(clarity="VS1", by=color)  
            , depth >= 1.2*do_by(depth, by=clarity, fun=mean, na.rm=TRUE)
                    )
out6 <- confront(diamonds, rules6)
summary(out6)



```




- Hierarchical aggregation structures: _hierarchy()_by


```{r, echo=TRUE, eval=FALSE}

#vha - example in our slides
## study first:
#?hierarchy

d <-  data.frame( code =c("01.1", "02.1", "01.2" 
                          ,"01.3", "02.2", "01", "02"
                          ,"A")
                , city = c("Seltjarnarnes", "Akureyri", "Reykjavik" 
                        ,"OtherCA", "OtherNonCA", "CapitalArea", "NonCA"
                        ,"All")
                , swimmingpools =c(1, 10, 1 ,5 , 22 ,7, 23, 30)
                  )

dgeo <- data.frame( 
                    code=c("A", "01", "02", "01.1", "01.2", "01.3", "02.1", "02.2")
                  , parent=c("-", "A", "A", "01", "01", "01", "02", "02")
                   )

d$check <- validate::hierarchy(d$swimmingpools, d$code, hierarchy = dgeo)

rule7 <- validator(
  hierarchy(swimmingpools, city, hierarchy=ref$codelist, by=list(city))
   )

# check 
# list(codelist=dgeo)$codelist
# to make sure 

out7 <- confront(d, rule7, ref=list(codelist=dgeo))
summary(out7)


```

May take a look at the examples in the handbook as well!


- Other checks made easy

We also learn how to use "check_that()", although we can reformulate by building validators


```{r, echo=TRUE, eval=FALSE}

# in_range

check_that(diamonds, in_range(z, 0, 32))
# is equivalent to
rule8 <- validator(in_range(z, 0, 32))
out8 <- confront(diamonds, rule8)
out8


```




```{r, echO=TRUE, eval=FALSE}
# contains_at_least
# does_not_contain
# contains_exactly


## Check that data is present for all years in 2000-2020
mydat <- data.frame(
      year = 2000:2020
    , value = sample(20:50,length(2000:2020))
    )

# Method 1: creating a data frame in-place (only for simple cases)
myrule <- validator(contains_exactly(
            year=c("2000","2010")
            ))
out <- confront(mydat, myrule)
out


# Method 2: pass the keyset to 'confront', and reference it in the rule.
# this scales to larger key sets but it needs a 'contract' between the
# rule definition and how 'confront' is called.
keyset <- c("2000","2010")
myrule2 <- validator(contains_exactly(all_keys))
out2 <- confront(mydat, myrule2, ref=list(all_keys = keyset))
out2

```


We could try more of the functions listed in chapter 7.6 at:

https://cran.r-project.org/web/packages/validate/vignettes/cookbook.html#26_Numeric_ranges


